### 2016-04-11 manuscript.tex
Fix errors from reflowing

% Template for PLoS
% Version 2.0 July 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
%
% -- FIGURES AND TABLES
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text. For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
%
% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
% For bolding characters in mathmode, please use $\mathbf{}$
%
% Please add line breaks to long equations when possible in order to fit our 2-column layout.
%
% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
%
% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
%
%
% % % % % % % % % % % % % % % % % % % % % % % %

@@

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

@@

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{????}
}
% Insert Author names, affiliations and corresponding author email.
\\
Mason Liang$^{1,\ast}$,
Rasmus Nielsen$^{2}$
\\
\bf{1} Mason Liang Integrative Biology, University of California, Berkeley, California, United States of America
\\
\bf{1} Rasmus Nielsen Integrative Biology, University of California, Berkeley, California, United States of America
\\
$\ast$ E-mail: wmliang@berkeley.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Estimating admixture histories is crucial for understanding the genetic diversity we see in present-day populations. Existing allele frequency or phylogeny-based methods are excellent for inferring the existence of admixture or its proportions, but have less power for estimating admixture times. Recently introduced approaches for estimating these times use spatial information from admixed chromosomes, such as the local ancestry or the decay of admixture linkage disequilibrium (ALD).  One popular method, implemented in the programs ALDER and ROLOFF, uses two-locus ALD to infer the time of a single admixture event, but is only able to estimate the time of the most recent admixture event based on this summary statistic.  We derive analytical expressions for the expected ALD in a three-locus system and provide a new statistical method based on these results that is able to resolve more complicated admixture histories. Using simulations, we show how this new statistic behaves on a range of admixture histories. As an example, we also apply our method to the Colombian and Mexican samples from the 1000 Genomes project.

\section*{Introduction}
There are many methods for inferring the presence of admixture, e.g. methods using simple summary statistics detecting deviations from phylogenetic symmetry \cite{reich2009reconstructing} [Also cite patterson paper and paper from Monty's group] , methods estimating admixture proportions with programs such as Structure \cite{pritchard2000inference},  Admixture \cite{alexander2009fast} or RFmic [cite] [comment: we might want to make sure to add more citations here, for example all the ones from our previous papers and the ones from the ALDER paper.  People like to see themselves cited]. However, there has been less research on estimating admixture times, possibly because such methods require data which were unavailable until the advent of high-throughput next generation sequencing. Some recently developed methods use the inferred local ancestry of sequences to construct admixture tract length distributions [cite Pool Nielsen, Gravel and yourself]. Over time, recombination is expected to decrease the average lengths of admixture tracts. The length distribution of admixture tracts is therefore informative about the time since admixture.  Much of the theory relating to tracts lengths is based on Fisher's famous theory of junctions \cite{fisher1949theory} and subsequent work by Stam \cite{stam1980distribution} and others [add as many citations here as possible]. For example, \cite{baird2003distribution} first discussed the length distribution of tracts descended from a single ancestor. These results informed later analyses of admixture tract length distribution, such as \cite{pool2009inference}, \cite{gravel2012population}, and \cite{liang2014lengths}. \cite{gravel2012population} also implemented the software program TRACTS, which estimates admixture histories by fitting the tract length distribution, obtained by local ancestry inference, to a exponential approximation.

@@

ROLLOFF and ALDER use the information contained in pairs of sites by examining the two-locus linkage disequilibrium between them. Here we extend the theory underlying the methods in ROLOFF and ADLER to three loci by considering three-locus LD. There are two ways of measuring the linkage between $n$ loci. \cite{bennett1952theory} defines $n$-locus linkage in a way that maintains a geometric decrease of LD each generation as a result of recombination, which is an important property of two-locus linkage disequilibrium. \cite{slatkin1972treating} defines $n$-locus LD  to be the $n$-way covariance, analogously to the property of two locus LD as the covariance in allele frequency between pairs of loci. For two and three loci, these two definitions coincide, but for four or more loci, they do not.

@@

[Could we say anything here about Neanderthals in Asia?]

% You may title this section "Methods" or "Models".
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"
\section{Model}
We use a random union of gametes admixture model as described in \cite{liang2014understanding}, which is an extension of the mechanistic admixture model formulated by \cite{verdu2011general}. In this model, two (or more) source populations contribute migrants to form an admixed population consisting of $2N$ haploid individuals. Each generation in the admixed population is formed through the recombination of randomly selected individuals from the previous generation, with some individuals potentially replaced by migrants from the source populations. For simplicity, we consider a model with only two source populations. Furthermore, the first source population only contributes migrants in the founding generation, $T$. The second source population contributes migrants in the founding generation and possibly in one or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before the present), a fraction $m_i$ of the admixed population is replaced by individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry}
ROLLOFF and ALDER use the standard two-locus measure of LD between a SNP at positions $x$ and another SNP at position $y$, which is a genetic distance $d$ to the right,
\begin{align}
D_2(d) = \text{cov}(H_x,H_y),
\label{D2}
\end{align}
where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$ if the $i^\text{th}$ sample is carrying the derived allele at the SNP at position $x$, and is otherwise 0. Alternatively, for genotype data, $H_{i,x}$ take on values from $\{0,1/2,1\}$ depending on the number of copies of the derived allele the $i^\text{th}$ sample is carrying at SNP position $x$.
We consider an additional site at position $z$, which is located a further genetic distance $d'$ to the right of $y$. The three-loci LD, as defined by as defined by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by

@@

When using only the admixed population itself as a reference population, the method described above will be biased if the same samples are used to estimate both the linkage disequilibrium coefficients and the weights ($\delta_x$, $\delta_y$, and $\delta_z$). We cannot efficiently compute a polyache statistics like \cite{loh2013inferring}. At the cost of some power, we instead adopt the approach of \cite{pickrell2012inference} and separate the admixed population into two equal-sized groups. We then use one group to estimate the weights, and the other group to estimate linkage disequilibrium coefficients, and vice versa. This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we then average. [Comment: I didn't understand how this works.  Coudl we five more detail?]
\subsection{Fitting the Two-Pulse Model}
We fit equation \ref{alder_2p} to the estimates of the weighted LD using non-linear least squares, with two modifications. We added a proportionality constant to account for the expected square allele frequency difference between the source populations. We also subtracted out an affine term in the weighted LD which is due to population substructure \cite{loh2013inferring}. We estimated this by computing the three-way covariance between triples of chromosomes. We use the jackknife to obtain confidence intervals for the resulting estimates by leaving out each chromosome in turn and refitting on the data for the remaining chromosomes.  [Comment: again, we need more detail here. These are important parts of the manuscript]
\section{Simulations}
We used the program macs \cite{chen2009fast} to generate two source populations which diverged 4000 generations ago and a coalescent simulation to generate an admixed population from the two source populations according to two-pulse and constant admixture models. We sampled 50 diploid individuals from the admixed and two source populations, each consisting of 20 chromosomes of length 1 Morgan. The effective population size was $2N=1000$ for the admixed population and two source populations. Using a two pulse model, we varied the migration probabilities and timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We also simulated data for a model with a constant rate of admixture each generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set}
We computed the weighted LD for the Mexican and Columbia populations in the first phase of the 1000 Genomes data set. These consisted of 66 individuals from Los Angeles and 60 individuals from Medellin, respectively. We used the 88 Yoruba samples as the one reference population. We computed the weighted LD on the genotypes to avoid effects of phasing errors.
% Results and Discussion can be combined.
\section{Results and Discussion}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\subsection{Simulations}

@@

[Comment: the color coding should be labeled for all figures.  Also, the y and x-axes should be labeled]

@@

[Comment: we need  figures showing the similar one-dimensional LD for comparison.  Maybe the first thing to discuss is to show how you can get approximately the same 1D curve for different models which produce different 2D surfaces]

@@

\begin{figure}
\includegraphics[scale=.6]{alder_Ts.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying pulse times. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Red lines are $T_2=2$, Green lines $T_2=5$, and blue lines are $T_2=10$.
}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{alder_ms.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying migration probabilities. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Increasing red values are higher $M_1$'s and increasing green values are higher $M_2$'s  (I will probably need to choose a better color scheme for this).
}
\end{figure}

>>>>

@@

% rotating package for sideways tables
% \usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below.
% Also, please see the algorithm section of our LaTeX guidelines
% (http://www.plosone.org/static/latexGuidelines) for important information about
% required formatting.
% \usepackage{algorithmic}
% \usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
% \usepackage{setspace}
% \doublespacing

@@

% Bold the 'Figure #' in the caption and separate it with a period % Captions
% will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

@@

% Title must be 150 characters or less
\begin{flushleft}
	{\Large \textbf{????} } \\

	%Insert Author names, affiliations and corresponding author email.
 	Mason Liang$^{1,\ast}$,
	Rasmus Nielsen$^{2}$\\

	\bf{1} Mason Liang Integrative Biology, University of California, Berkeley, California, United States of America\\
	\bf{1} Rasmus Nielsen Integrative Biology, University of California, Berkeley, California, United States of America\\
	$\ast$ E-mail: wmliang@berkeley.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Estimating admixture histories is crucial for understanding the genetic
diversity we see in present-day populations. Existing allele frequency or
phylogeny-based methods are excellent for inferring the existence of admixture
or its proportions, but have less power for estimating admixture times. Recently
introduced approaches for estimating these times use spatial information from
admixed chromosomes, such as the local ancestry or the decay of admixture
linkage disequilibrium (ALD).  One popular method, implemented in the programs
ALDER and ROLLOFF, uses two-locus ALD to infer the time of a single admixture
event, but is only able to estimate the time of the most recent admixture event
based on this summary statistic.  We derive analytical expressions for the
expected ALD in a three-locus system and provide a new statistical method based
on these results that is able to resolve more complicated admixture histories.
Using simulations, we show how this new statistic behaves on a range of
admixture histories. As an example, we also apply our method to the Colombian
and Mexican samples from the 1000 Genomes project.

\section*{Introduction}
There are many methods for inferring the presence of
admixture, e.g. methods using simple summary statistics detecting deviations
from phylogenetic symmetry \cite{reich2009reconstructing} [TODO(mason) Also cite
patterson paper and paper from Monty's group], methods estimating admixture
proportions with programs such as Structure \cite{pritchard2000inference},
Admixture \cite{alexander2009fast} or RFmix \cite{maples2013rfmix}. However,
there has been less research on estimating admixture times,
possibly because such methods require data which were unavailable until the
advent of high-throughput next generation sequencing. Some recently developed
methods use the inferred local ancestry of sequences to construct admixture
tract length distributions, such as \cite{pool2009inference},
\cite{gravel2012population}, and \cite{liang2014lengths}.
Over time, recombination is expected to decrease the average lengths of admixture
tracts. The length distribution of admixture tracts is therefore informative
about the time since admixture.  Much of the theory relating to tracts lengths
is based on Fisher's famous theory of junctions \cite{fisher1949theory} and
subsequent work, such as
\cite{
	stam1980distribution,
	guo1994computation,
	bickeboller1996distribution,
	bickeboller1996probability,
	stefanov2000distribution,
	ball2005evaluation,
	cannings2003identity,
	dimitropoulou2003recsim,
	walters2005probability,
	rodolphe2008theoretical
}.
For example, \cite{baird2003distribution} first
discussed the length distribution of tracts descended from a single ancestor.
These results informed later analyses of admixture tract length distribution,
such as \cite{pool2009inference}, \cite{gravel2012population}, and
\cite{liang2014lengths}. \cite{gravel2012population} also implemented the
software program TRACTS, which estimates admixture histories by fitting the
tract length distribution, obtained by local ancestry inference, to a
exponential approximation.

@@

ROLLOFF and ALDER use the information contained in pairs of sites by examining
the two-locus linkage disequilibrium between them. Here we extend the theory
underlying the methods in ROLLOFF and ADLER to three loci by considering
three-locus LD. There are two ways of measuring the linkage between $n$ loci.
\cite{bennett1952theory} defines $n$-locus linkage in a way that maintains a
geometric decrease of LD each generation as a result of recombination, which is
an important property of two-locus linkage disequilibrium.
\cite{slatkin1972treating} defines $n$-locus LD  to be the $n$-way covariance,
analogously to the property of two locus LD as the covariance in allele
frequency between pairs of loci. For two and three loci, these two definitions
coincide, but for four or more loci, they do not.

@@

[TODO Could we say anything here about Neanderthals in Asia?]

% You may title this section "Methods" or "Models". % "Models" is not a valid
title for PLoS ONE authors. However, PLoS ONE % authors may use "Analysis"
\section{Model} We use a random union of gametes admixture model as described in
\cite{liang2014understanding}, which is an extension of the mechanistic
admixture model formulated by \cite{verdu2011general}. In this model, two or
more source populations contribute migrants to form an admixed population
consisting of $2N$ haploid individuals. Each generation in the admixed
population is formed through the recombination of randomly selected individuals
from the previous generation, with some individuals potentially replaced by
migrants from the source populations. For simplicity, we consider a model with
only two source populations. Furthermore, the first source population only
contributes migrants in the founding generation, $T$. The second source
population contributes migrants in the founding generation and possibly in one
or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before
the present), a fraction $m_i$ of the admixed population is replaced by
individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry} ROLLOFF and ALDER use the
standard two-locus measure of LD between a SNP at positions $x$ and another SNP
at position $y$, which is a genetic distance $d$ to the right,

\begin{align} D_2(d) = \text{cov}(H_x,H_y), \label{D2} \end{align}

where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed
chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$
if the $i^\text{th}$ sample is carrying the derived allele at the SNP at
position $x$, and is otherwise 0. Alternatively, for genotype data, $H_{i,x}$
take on values from $\{0,1/2,1\}$ depending on the number of copies of the
derived allele the $i^\text{th}$ sample is carrying at SNP position $x$. We
consider an additional site at position $z$, which is located a further genetic
distance $d'$ to the right of $y$. The three-loci LD, as defined by as defined
by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by

@@

When using only the admixed population itself as a reference population, the
method described above will be biased if the same samples are used to estimate
both the linkage disequilibrium coefficients and the weights ($\delta_x$,
$\delta_y$, and $\delta_z$). We cannot efficiently compute a polyache statistics
like \cite{loh2013inferring}. At the cost of some power, we instead adopt the
approach of \cite{pickrell2012inference} and separate the admixed population
into two equal-sized groups. We then use one group to estimate the weights, and
the other group to estimate linkage disequilibrium coefficients, and vice versa.
This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we
then average. [TODO Comment: I didn't understand how this works.  Coudl we five
more detail?] \subsection{Fitting the Two-Pulse Model} We fit equation
\ref{alder_2p} to the estimates of the weighted LD using non-linear least
squares, with two modifications. We added a proportionality constant to account
for the expected square allele frequency difference between the source
populations. We also subtracted out an affine term in the weighted LD which is
due to population substructure \cite{loh2013inferring}. We estimated this by
computing the three-way covariance between triples of chromosomes. We use the
jackknife to obtain confidence intervals for the resulting estimates by leaving
out each chromosome in turn and refitting on the data for the remaining
chromosomes.  [TODO Comment: again, we need more detail here. These are
important parts of the manuscript] \section{Simulations} We used the program
macs \cite{chen2009fast} to generate two source populations which diverged 4000
generations ago and a coalescent simulation to generate an admixed population
from the two source populations according to two-pulse and constant admixture
models. We sampled 50 diploid individuals from the admixed and two source
populations, each consisting of 20 chromosomes of length 1 Morgan. The effective
population size was $2N=1000$ for the admixed population and two source
populations. Using a two pulse model, we varied the migration probabilities and
timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We
also simulated data for a model with a constant rate of admixture each
generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set} We computed the weighted LD for the Mexican and Columbia
populations in the first phase of the 1000 Genomes data set. These consisted of
66 individuals from Los Angeles and 60 individuals from Medellin, respectively.
We used the 88 Yoruba samples as the one reference population. We computed the
weighted LD on the genotypes to avoid effects of phasing errors. % Results and
Discussion can be combined. \section{Results and Discussion} % We only support
three levels of headings, please do not create a heading level below
\subsubsection. \subsection{Simulations}

@@

[TODO: the color coding should be labeled for all figures.  Also, the y and
x-axes should be labeled]

@@

[TODO Comment: we need  figures showing the similar one-dimensional LD for
comparison.  Maybe the first thing to discuss is to show how you can get
approximately the same 1D curve for different models which produce different 2D
surfaces]

@@

\begin{figure}
	\includegraphics[scale=.6]{alder_Ts.eps}
	\caption{
		{\bf Extra figure}
		Corresponding Alder curves for two-pulse admixture with varying pulse times.
		Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Red lines are
		$T_2=2$, Green lines $T_2=5$, and blue lines are $T_2=10$.
	}
\end{figure}

\begin{figure}
	\includegraphics[scale=.6]{alder_ms.eps}
	\caption{
		{\bf Extra figure}
		Corresponding alder curves for two-pulse admixture with varying migration
		probabilities. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis.
		Increasing red values are higher $M_1$'s and increasing green values are
		higher $M_2$'s  (TODO: choose better colors).
	}
\end{figure}
